Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-08 p2/def-commands.tex
Improvements to the T mapping

For example, $\cbfaa{n}$ will represent $create(n,\valf)$,
and $\cdba{n}$ will represent $rmdir(n)$.
When describing unknown commands, we substitute one or both
of their values with variables, and write e.g. $\cxynv$.
By $\cxynv=\czwmv$ we mean that the commands are identical
($[x]=[z]$ and $y=z$ and $n=m$).
As suggested above, we will frequently only investigate
the types of the input and output values of a command
(its input and output types), and write
$\cxy=\czw$, by which we mean $[x]=[z]$ and $[y]=[w]$.

>>>>

For example, $\cbfaa{n}$ will represent $create(n,\valf)$,
and $\cdba{n}$ will represent $rmdir(n)$.
When describing unknown commands, we substitute one or both
of their values with variables, and write e.g. $\cxynv$.
By $\cxynv=\czwmv$ we mean that the commands are identical
($[x]=[z]$ and $y=z$ and $n=m$).
As suggested above, we will frequently only investigate
the types of the input and output values of a command
(its input and output types), and write
$\cxy=\czw$, by which we mean $[x]=[z]$ and $[y]=[w]$
(or equivalently $x\typeeq z$ and $y\typeeq w$).

### 2016-10-08 p2/works-inverse.tex
Improvements to the T mapping

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
%% TODO extending it!
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems we write $\FS\typeeq\GS$ iff they are both broken, or they share the same nodes $\setn$,
and for all $n\in\setn$, $[\FS(n)]=[\GS(n)]$.
\item For two commands we write $\cxynv\typeeq\czwnv$ iff 
$[x]=[z]$ and $[y]=[w]$ and $n=m$.
\item For two sequences of commands we write $A\typeeq B$ iff they contain the same number of commands,
and pairwise the commands $\alpha\in A$ and $\beta\in B$ at the same indices are equivalent by type ($\alpha\typeeq\beta$).
\end{itemize}
\end{mydef}

In other words, two type-equivalent objects can only differ in the value they contain,
but as $|\setb|=|\setd|=1$, this is only possible at values of the type $[\valf]$.
We select an arbitrary file value $\fil\in[\valf]$, and choose a representative
of each $\typeeq$-equivalence class where all (if any) values of the type $[\valf]$ is $\fil$.
For example, in the case of commands, the representative of $[\cbfaa{n}]_\typeeq$ is $\caaa{\empt}{\fil}{n}$.

\begin{mydef}[$\T$]
We use $\T$ and write $X\T$ to denote the canonical mapping of 
filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
$\T$ is also extended to sets of sequences,
where $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{mydef}

For example, $\cbfaa{n}\T = \caaa{\empt}{\fil}{n}$,
and for the sequence $S=\alpha\cc\beta$, $S\T=\alpha\T\cc\beta\T$.
In other words, $\T$ maps our system of filesystems and commands
into a system where $|\setf|$ is also $1$.

@@

\begin{mylem}
Representative commands $\alpha\T$ are injections over representative filesystems ($\FS\T$)
except for the cases where they break a filesystem, that is,
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying the command to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
We know that a representative filesystem can only have values from the set $\rvalueset$.
As $\alpha\T(\FS\T)\neq\fsbroken$, $\FS\T(n)$ must be the value that matches the input type of $\alpha\T$.
We know that $\GS\T(n)\neq\FS\T(n)$, and therefore $\GS\T(n)$ must have a different type,
which implies $\alpha\T(\GS\T)=\fsbroken$ as it cannot match the input type of $\alpha\T$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

\begin{mydef}[Inverse commands and sequences]
If we regard the representative command $\alpha\T$ a function whose
domain is the subset of representative filesystems $\{\FS\T | \alpha\aFS\neq\fsbroken\}$,
then we can define the inverse of $\alpha\T$
for which we write $\alpha\T^{-1}$.
We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}

%% TODO
% Do we need <[x],[y],n>-1 = <[y],[x],n> ?
% Do we need to prove if a filesystem is not in the image, then a-1 will break it?

From the definitions we can clearly see that
\begin{mycor}\label{negneg_is_typeeq} %% TODO Don't use!
$\forall S, \FS: (S^{-1})^{-1}\aFS\typeeq S\aFS$.
\end{mycor}

>>>>

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
%% TODO extending it!
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems $\FS\typeeq\GS$ iff they are both broken, or they share the same nodes $\setn$,
and for all $n\in\setn$, $\FS(n) \typeeq \GS(n)$.
\item For two commands $\cxynv\typeeq\czwnv$ iff 
$x\typeeq z$ and $y\typeeq w$ and $n=m$.
\item For two sequences of commands $\emptyseq\typeeq\emptyseq$, and
if $A'=\alpha\cc A$ and $B'=\beta\cc B$, then $A'\typeeq B'$ iff $\alpha\typeeq\beta$ and $A\typeeq B$.
\end{itemize}
\end{mydef}

We select representatives of each $\typeeq$-equivalence class of values:
for $[\empt]$ and $[\vald]$ this must be $\empt$ and $\vald$; and
for $[\valf]$ we select an arbitrary file value $\fil\in[\valf]$.
We then use these representative values to
select representatives of the equivalence classes of
other objects in the following way:

\begin{mydef}[$\T$]
We use $\T$ and write $X\T$ to denote the canonical mapping of 
values, filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
\begin{itemize}
\item For values, $\vald\T=\vald$, $\valf\T=\fil$ and $\empt\T=\empt$.
\item For filesystems, $\FS\T$ is defined by $\FS\T(n) = \FS(n)\T$.
\item For commands, $\cxynv\T = \cTxynv$.
\item For sequences of commands, $\emptyseq\T=\emptyseq$, and if $S'=\alpha\cc S$, then $S'\T=\alpha\T\cc S\T$.
\item Finally we extend $\T$ to sets of sequences: $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{itemize}
\end{mydef}

It is easy to see that for all objects where it is defined, $X\T\typeeq X$.

In other words, $\T$ maps our system of filesystems and commands
into a \emph{representative} system where there is a one-to-one correspondence between types
and values.
% $|\setf|$ is also $1$.
At the same time, representative filesystems, commands, sequences, etc.
are subsets of their non-representative counterparts; hence all
propositions proven so far apply to them as well.

@@

\begin{mylem}\label{repr_comm_inject}
Representative commands $\alpha\T$ are injections over representative filesystems ($\FS\T$)
except for the cases where they break a filesystem, that is,
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
Let $\alpha\T = \cTxynv$.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying $\cTxynv$ to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
As $\cTxynv(\FS\T)\neq\fsbroken$, we know $[\FS\T(n)]=[x]$.
We also know that $\GS\T(n)\neq\FS\T(n)$, and therefore, as there is a one-to-one correspondence between values and types,
$[\GS\T(n)]\neq[x]$.
This implies $\cTxynv(\GS\T)=\fsbroken$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

\begin{mydef}[Inverse commands and sequences]
If based on \cref{repr_comm_inject}
we regard the representative command $\alpha\T$ a bijective function whose
domain is the subset of representative filesystems $\{\FS\T | \alpha\aFS\neq\fsbroken\}$,
then we can define the inverse of $\alpha\T$
for which we write $\alpha\T^{-1}$.
For any $\GS\T$ outside the range of $\alpha\T$ we define $\alpha\T^{-1}(\GS\T)$ to be $\fsbroken$.

We write ${S\T}^{-1}$ for the inverse of the sequence $S\T$, which consists of the inverses of the commands in $S\T$
in reverse order.
\end{mydef}

%% TODO
% Do we need <[x],[y],n>-1 = <[y],[x],n> ?
% Do we need to prove if a filesystem is not in the image, then a-1 will break it?
\begin{mylem}
For representative commands
\[ \cTxynv^{-1} = \cTyxnv \]
\end{mylem}
TODO

From the definitions we can clearly see that
\begin{mycor}\label{negneg_is_typeeq}
$\forall S: ({S\T}^{-1})^{-1} = S\T$.
\end{mycor}
